Cover letter for the original IJRR submission

### 2017-10-10 reviews_tro/reviewer_response.tex

\quot { (1) A criterion for robust static equilibrium, but it isn't
described until page 12, in Appendix B, Section B. [\dots]  It would really help to describe this MUCH earlier in
the paper (than page 13), if it is a significant novelty of your paper.}

\as{We agree with the remark. To better distinguish between what is novel and not, we modified the paper to move the content of Appendix B.B (now Section VI), as well as D  (the pseudo code of our algorithm, now Section V.B.2) and E (reference to our source code, now Section VII) to the text itself. This results in the criterion being presented much earlier.}\done \jp

\quot{It is a pretty intuitive/typical check for quantifying a
margin for friction conesque violations (using an approximation for
the friction cone, with x and y direction forces considered
independently). }

\as{Thanks for the comment. To our knowledge, what is common in the literature is to reduce the friction coefficient to obtain more conservative bounds on the admissible forces, and thus a ``safety margin'' to the real cone boundary. An issue with this formulation is that the resulting margin may be too small for small forces, and too large for large forces. With our formulation instead, the margin is constant (i.e. the value $b_0$). We updated the paper to explain this difference. We believe this formulation is novel, but would be happy to cite any relevant reference if we happen to be wrong. This formulation proves useful to us because the margin $b_0$ allows to choose the more robust candidate for static equilibrium. }\done
\qt{(Section VI.2) A simple approach to robustness consists in choosing a smaller friction coefficient, to constrain the forces to lie away from the boundaries of the real cone.  However, this would result in a small safety margin for forces of low magnitude, and an excessively large safety margin for large forces as the boundaries grow more and more apart. In comparison, our margin $b_0$ is constant, and provides a helpful mean to compare the robustness of different contact configurations.}

### 2017-10-10 body/intro.tex
typos
Modif contribs.

Our solution belongs to the class of decoupled approaches, 
i.e. 
we
propose
specific
algorithms
to
efficiently
solve
both
\Pa
and \Pb
while 
relying 
on 
state-of-the-art 
solution 
to 
$\mathcal{P}_3$
to 
obtain
the
whole
movement.
Our
main
contribution
is
the
definition        
of
a reduction property, the reachability condition.

%~ because we believe it is the most promising direction~\citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of acyclic contact planning. 
Compared to previous approaches, our solution has two main novelties:

\noindent \textbf{Regarding $\mathcal{P}_1$}. We propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low-dimensional space for which there exists a mapping to the contact manifold.

@@

The reachability condition is the key to the strict separation between $\mathcal{P}_1$ and $\mathcal{P}_2$, hence to the low complexity of our planner, though it
can result in failures. However, we demonstrate empirically its interest: the high success rate and low computation times allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

To further demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times to address sequentially the three problems, and compare them to related work, demonstrating that our method is orders of magnitude faster.

Finally, we provide an extensive discussion on the consequences of our approach in terms of efficiency and completeness regarding the contact planning problem. \\

\noindent \textbf{Comparison with our previous work:}
The present paper is an extension of our ISRR conference paper~\cite{tonneauisrr15}. 
As such, our solution to address $\mathcal{P}_1$ and $\mathcal{P}_2$ is the same motion planner as the one presented at ISRR (reformulated in Section~\ref{rbprm} and~\ref{sec:contact}). 
However three important novelties have been added to the planner: the pseudo-code of the algorithm (Section~\ref{app:contact}), a novel criterion for
static equilibrium (Section~\ref{sec:Equil}), and the release of the source code of the planner (Section~\ref{app:hpp}).

The other contribution of the paper is a rigorous experimental validation of the approach on actual robot models (Section~\ref{sec:results}).
To validate our contact plans we introduced a complete framework for multi-contact motion synthesis. This framework additionally comprises an interpolation method to solve the problem $\mathcal{P}_3$,  based on a reformulation of our previous work~\cite{Carpentier2016}. Our solution to $\mathcal{P}_3$ allows us to verify that the synthesized motions are physically consistent, using our implementation of a state-of-the-art simulation algorithm~\cite{Kaufman2008}. 
We do not consider these additions as contributions \textit{per se}, but describing them here is required to understand our experimental contribution.

Finally, as opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are more constrained kinematically than the virtual avatars used before,
thus making the planning problem harder. Through an extensive empirical analysis, we demonstrate that the method remains efficient and applicable in this context.
%~ These aspects of the framework are presented in details in the paper, but are not novel \textit{per se}. The novelty lies in the validation of the method with real robot models, for which the planning is harder with respect to the avatars used in [20], because of stricter kinematic constraints

>>>>

Our solution belongs to the class of decoupled approaches, 
i.e. 
we
propose
specific
algorithms
to
efficiently
solve
both
\Pa
and \Pb
while 
relying 
on 
state-of-the-art 
solution 
to 
$\mathcal{P}_3$
to 
obtain
the
whole
movement.
\textbf{Our
main
contribution
is
the
definition        
of
a reduction property, the reachability condition.
}
%~ because we believe it is the most promising direction~\citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of acyclic contact planning. 
Compared to previous approaches of the litterature, our solution has two main novelties:

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low-dimensional space for which there exists a mapping to the contact manifold.

@@

\textbf{The reachability condition is the key to the strict separation between $\mathcal{P}_1$ and $\mathcal{P}_2$}, hence to the low complexity of our planner.
However, the reduction may result in failures. We demonstrate empirically its interest, both in the high success rate and in the low computation times, by a deep experimental validation, with two realistic robot models (humanoid and quadruped) in various scenarios, involving statistics of the relevance of our hypotheses.
These two properties allow us to plan, and re-plan upon failure, multi-contact sequences at \gls{interactive}$^1$ rates.

%% To further demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times to address sequentially the three problems, and compare them to related work, demonstrating that our method is orders of magnitude faster.

%% Finally, we provide an extensive discussion on the consequences of our approach in terms of efficiency and completeness regarding the contact planning problem. \\

%\noindent \textbf{Comparison with our previous work:}
\subsection{Organization and comparison with preliminary paper~\cite{tonneauisrr15}}
The present paper is an extension of our ISRR conference paper~\cite{tonneauisrr15}, that it aims at replacing.
The core methodological part corresponds to the same motion planner, with important complements listed below.
The present paper also proposes a novel experimental section, with statistical validation of the planner hypothesis using realistic robot models.

After presenting an overview of our method (Sec.~\ref{overview}) and introducing our notations (Sec.~\ref{notations}),
we present our solution to address $\mathcal{P}_1$ (Sec.~\ref{rbprm}) and $\mathcal{P}_2$ (Sec.~\ref{sec:contact}) , which corresponds to the same motion planner as the one presented at \cite{tonneauisrr15}. The two sections hence are reformulation of the preliminary paper. 
Three important novelties have been added to the planner: the pseudo-code of the algorithm (Sec.~\ref{app:contact}), a novel criterion for
static equilibrium (Sec.~\ref{sec:Equil}), and the release of the source code of the planner (Sec.~\ref{app:hpp}).

The other novelty of the paper is a rigorous experimental validation of the approach on actual robot models (Sec.~\ref{sec:results}).
We had to introduce a complete framework for multi-contact motion synthesis, in order to validate our contact plans. It comprises an interpolation method to solve the problem $\mathcal{P}_3$, based on a reformulation of our previous work~\cite{Carpentier2016}. The generated motions are verified to be physically consistent, using our implementation of a state-of-the-art simulation algorithm~\cite{Kaufman2008}. 
These aspects of the framework are presented in details in the paper, but are not novel \textit{per se}. The novelty lies in the complete validation of the contact planner with real robot models, for which the planning is much harder with respect to the avatars used in \cite{tonneauisrr15}, because of stricter kinematic constraints.

%We do not consider these additions as contributions \textit{per se}, but describing them here is required to understand our experimental contribution.

%Finally, as opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are more constrained kinematically than the virtual avatars used before, thus making the planning problem harder. Through an extensive empirical analysis, we demonstrate that the method remains efficient and applicable in this context.

%~ These aspects of the framework are presented in details in the paper, but are not novel \textit{per se}. The novelty lies in the validation of the method with real robot models, for which the planning is harder with respect to the avatars used in [20], because of stricter kinematic constraints
